---
title: "dplyr-101"
author: "Andres Fiallos"
format: html
---


```{r}
#| message: false
here::i_am("git-101-2024.Rproj")
library(vroom)
library(here)
library(ggplot2)
library(readr)
library(dplyr)
```
## Data import

```{r}
white_wine <- read.csv("~/Documents/QE_dauphine_m1/data_management/R/git-101-2024/white_wine.csv", sep=";")
```



|     |     |
|-----|-----|
|  Number of observations   |  `r NROW(white_wine)`   |
|  Number of variables   |  `r NCOL(white_wine)`   |


This piece of code does not work due to `quality` being a numeric variable which does not allow for filling
```{r}
#| eval: false
#| code-fold: false
ggplot(white_wine, aes(x = quality, fill = quality))+
  geom_bar()
```



## Data transformation

```{r}
#| message: false
library(dplyr)


```

### Column extraction

```{r}
class(white_wine)

class(pull(white_wine, quality))
```

With `dplyr`, `pull()` is the column extraction function, somewhat similiar to the dollar `$` operation and to the the double bracket `[[]]`.

### Column recoding
```{r}
#| eval: false

white_wine$quality <- factor(white_wine$quality, ordered = T, levels = 3:9)
white_wine$quality <- white_wine$quality |> factor(ordered = T, levels = 3:9)
```

```{r}
white_wine <- white_wine |> mutate(quality = factor(quality,ordered = T, levels = 3:9))

ggplot(white_wine, aes(x = quality, fill = quality))+
  geom_bar()
```

